---
title: "Apple Quality Data Visualization"
author: "Kevin Pham"
date: "2024-03-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
apple_data <- read.csv("apple_quality.csv")
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(ggthemes)
library(gplots)
```

## R Markdown
The intended purpose of this document will be to demonstrate my familiarity with R's most widely used data visualization library ggplot2. Practice makes perfect. 

The following data set was retrieved from https://www.kaggle.com/datasets/nelgiriyewithana/apple-quality/data

Laughter is the best medicine, and an apple a day keeps the doctor away!

```{r quick summary}
summary(apple_data)
```
\newpage
## Cleaning the data


```{r cleaning}
na_count <- colSums(is.na(apple_data))
new_apple_data<- na.omit(apple_data)
good_apples <- new_apple_data %>% filter(Quality == "good") # Good apples
bad_apples <- new_apple_data %>% filter(Quality != "good") # Bad apples!
new_apple_data$Acidity <- as.numeric(new_apple_data$Acidity)
```

## Plots

This scatter plot shows the relationship between the scores assigned to the size and weight of apples in the dataset. Each point on the plot represents an individual apple, The x-axis indicates the Size Score and the y-axis represents the Weight Score. The color of each point is determined by the combined score of size and weight, creating a gradient from red to green.

```{r plotting}
ggplot(new_apple_data, aes(x = Size, y = Weight, color = Size + Weight)) +
  geom_point() +
  labs(title = "Scatter Plot of Apple Size vs. Weight Scores",
       x = "Size Score", y = "Weight Score") +
  scale_color_gradient(low = "red", high = "green") +
  theme(plot.title = element_text(hjust = 0.5))
```
This is a plot for ALL apples.
I would like to see if a "Good" apple is characterized by having higher weight and size scores.

## Good apples
```{r, good apples}
ggplot(good_apples, aes(x = Size, y = Weight, color = Size + Weight)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Scatter Plot of GOOD Apples Size vs. Weight Scores",
       x = "Size Score", y = "Weight Score") +
  scale_color_gradient(low = "red", high = "green") +
  theme(plot.title = element_text(hjust = 0.5))
```
## Bad apples
```{r bad apples}
ggplot(bad_apples, aes(x = Size, y = Weight, color = Size + Weight)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Scatter Plot of BAD Apples Size vs. Weight Scores",
       x = "Size Score", y = "Weight Score") +
  scale_color_gradient(low = "red", high = "green") +
  theme(plot.title = element_text(hjust = 0.5))
```
## Conclusion 1


* The negative correlation coefficient (-0.1707017) indicates a weak negative linear relationship between the "Weight" and "Size" variables.
* The test statistic (t = -10.954) is far from zero.
* The p-value is extremely small, suggesting strong evidence against the null hypothesis of no correlation.
* The 95 percent confidence interval does not include zero, further supporting the rejection of the null hypothesis.
* There is statistically significant evidence of a weak negative linear correlation between the "Weight" and "Size" variables in the dataset.

```{r correlation}
cor_test_result <- cor.test(new_apple_data$Weight, new_apple_data$Size, method = "pearson")
```

## More Correlations

Juiciness and Acidity are positively correlated. 
```{r}
cor_matrix <- cor(new_apple_data[, c("Weight", "Sweetness", "Crunchiness", "Juiciness", "Ripeness", "Acidity")])

# Create a heatmap of the correlation matrix with annotations and color map
heatmap.2(
  cor_matrix, 
  trace = "none", 
  col = colorRampPalette(c("blue", "white", "red"))(100), 
  margins = c(5, 10), 
  dendrogram = "none", 
  notecol = "black", 
  density.info = "none", 
  main = "Correlation Heatmap", 
  cexRow = 1,  
  cexCol = 1   
)
```

